---
title: "Mall Customers Capstone"
author: "Samantha Jones"
date: "2024-10-07"
output: html_document
---
##
Introduction: An owner of a supermarket mall with active customer memberships wants to analyze some basic customers information from their  memberships such as:Customer Id,Gender,Age,Annual Income,Spending Score. Spending Score is assigned to customers based on the owner perimeters of customer behavior and purchasing data.


Problem: Owner wants to understand who their Target Audience is? For their marketing team to increase their promotional strategies.

```{r}
library(readr)
Mall_Customers <- read_csv("Mall_Customers.csv")
summary(Mall_Customers)


```{r}
library(tidyverse)
library(ggplot2)
library(conflicted)
library(dbplyr)
```
Check for missing values:
```{r}
sum(is.na(Mall_Customers))
```
Check for duplicates:
```{r}
duplicated_rows <- duplicated(Mall_Customers)
sum(duplicated_rows)
```
Rename columns for clarity:
```{r}
names(Mall_Customers) <- c("CustomerID", "Gender", "Age", "Annual_Income", "Spending_Score")
```
Convert the annual income from double digits $15 and currency K$ into thousands as this will make for a more accurate annual income column in US dollars: 
```{r}
Mall_Customers$Annual_Income <- (Mall_Customers$Annual_Income * 1000)
summary(Mall_Customers)
```
Data type conversion:
```{r}
Mall_Customers$Annual_Income <- as.numeric(Mall_Customers$Annual_Income)
Mall_Customers$Spending_Score <- as.numeric(Mall_Customers$Spending_Score)
```
Segmentation by Gender:
```{r}
aggregate(cbind(Annual_Income, Spending_Score)~ Gender, data = Mall_Customers, FUN = mean)
```
As you can see even though Men make almost about 3000 dollars more than Women spend about 3% according to Spending Score.


Customer Age Grouping: 
Create age groups to study patterns across age brackets.
```{r}
Mall_Customers$Age_Group <- cut(Mall_Customers$Age, 
                                breaks = c(0, 20, 40, 60, 80), labels = c("0-20", "21-40", "41-60", "61-80"))
aggregate(Spending_Score ~ Age_Group, data = Mall_Customers, FUN = mean)
```
As you can see from the data table the age groups that spend the most range from ages 21-40. 


Annual Income Analysis:
Divide income into categories to study correlations with spending. 
```{r}
Mall_Customers$Income_Group <- cut(Mall_Customers$Annual_Income, breaks = 4, labels = c("Low", "Mid", "High", "Very High"))
aggregate(Spending_Score ~ Income_Group, data = Mall_Customers, FUN = mean)
```
```
SPECIFIC STEPS IN THE ANALYTICS PROCESS
MARKET RESEARCH:

1.Define objectives: understand the business needs to identify customers segments based on demographics and spending behavior.


2.Data Exploration: Conducted exploratory analysis(EDA) to summarize main characteristics using statistical graphics and visulation(BELOW).


3.Segmentation: Use clustering or catergorization techniques to identify different customer segments based on the factors such as income,age, and gender.


4. Insights & Recommendations:Derive actionable insights for retail marketing strategies, such as targeted campaigns for specific age or income groups.

GUIDE TO VISUALIZATION:
OBJECTIVE:Present my findings visually to communicate key insights: 

Age Distribution:
```{r}
library(ggplot2)
```
```{r}
ggplot(Mall_Customers, aes(x = Age_Group)) + geom_bar(fill = "blue") + ggtitle("Distribution of Customers by Age")
```


Analysis: What age group spends the most?


Age group that shops the most is ranging between 21-40


Spending Score by Gender: 
```{r}
ggplot(Mall_Customers, aes(x = Gender, y = Spending_Score, fill = Gender)) + geom_boxplot() + ggtitle("Spending Score by Gender")
```

Analysis: Which gender has a higher Spending Score?


Data shows that women have a higher Spending Score then men

Annual Income vs. Spending Score:
```{r}
ggplot(Mall_Customers, aes(x = Annual_Income, y = Spending_Score, color = Gender)) + geom_point() + ggtitle("Annual Income vs. Spending Score by Gender")
```

Analyis: Is there a positive correlation between income and spending habits? Does Gender influence this?


The bulk of both genders that make between 40000-70000 have just about the same spending habits. I don't believe gender is a major influnce on this because genders that make above 100k spend either less or more, those making below 50k have a low or high spending score as well.



Income Group vs. Spending Score:
```{r}
ggplot(Mall_Customers, aes(x = Income_Group, y = Spending_Score, fill = Income_Group)) + geom_boxplot() + ggtitle("Spending Score by Income Group")
```


Analysis: Does a higher income group translate into higher spending?

Low Range Income:15000-44000

Mid Range Income:46000-76000

High Range Income:77000-103000

Very High Range Income: 113000-137000


Mid range income spending score falls within the 50 range, shoppers either below or above the average range


Low range income spending score is very consistent within the 25-75 range


High range income spending score is within the 15-80 range but overall is consistent


Very high range income spending score is within the 15-75 range is consistent as well



CONCLUSION: 

From this dataset I've concluded that age does factor into customers spending habits, moreso customers from ages 21-40. Income is generally not something that is keeping shoppers from spending, but 3 out of 4 income brackets(Low,High,VeryHigh) stay within reason. On the other hand mid income range does differ, meaning they spend to much, likely not at all or pretty adverage.


Both male and female spending habits were just about the same, women just slight spend a little more than men.

To further dive into this dataset maybe a catagory for items customers shop for i.e.(Electronic,Groceries,Clothing,Travel) would have helped to see where customers spend the most frequent.I believe if that data was present the marketing team could run sells/ads more geared towards those items.

However since the target age audience is 21-40, advertisement on social media would be very benefical to the owner. 


```

```


